\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry} 
\usepackage{amsmath,amsthm,amssymb,amsfonts}
\usepackage{hyperref,enumerate}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}
\urlstyle{same} 

\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
 
\newenvironment{problem}[2][Problem:]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}\hskip \labelsep {\bfseries #2.}]}{\end{trivlist}}
%If you want to title your bold things something different just make another thing exactly like this but replace "problem" with the name of the thing you want, like theorem or lemma or whatever
 
\begin{document}
 
%\renewcommand{\qedsymbol}{\filledbox}
%Good resources for looking up how to do stuff:
%Binary operators: http://www.access2science.com/latex/Binary.html
%General help: http://en.wikibooks.org/wiki/LaTeX/Mathematics
%Or just google stuff


\title{ENGS/QBS 108 Fall 2017 Assignment 4 Part 1}
\author{Due October 31, 2017 \\ Instructors: George Cybenko and Saeed Hassanpour \\ Prepared by: Benjamin Priest}
\date{}
\maketitle

\pagebreak

%\section{$K$ means clustering and $K$ nearest neighbors classification}
\begin{problem}{Natural Language Processing [50 points]}
In this problem, you will exercise your knowledge of natural language processing by way of a few problems.
Further problems concerning reinforcement learning will be assigned later in the week. 

\begin{enumerate}
    \item {[20 points]} 
    We will begin by building a language model utilizing 
    \href{https://canvas.dartmouth.edu/files/3206548/download?download_frd=1}{War and Peace}.
    \begin{enumerate}
            \item Change all alphabetic characters to lowercase and all non-alphabetic characters in the text to new-lines 
            	(If you have time, read the novel first, before changing it).
	\item Tokenize the text file and output the first 10 words of the book.
	\item Sort the words alphabetically and output the first 10 words of the book.
	\item Count the unique words in the book and output the first 10 (unique) words in the alphabetical order and their counts.
	\item What are the 10 most frequent words in the book and what are their frequencies?
    \end{enumerate}

    \item {[15 points]}
    Calculate the minimum edits distance between these pairs of 
    \href{http://www.barbapapa.com/the-barbapa-family-en/}{Barbapapa family} members? names. 
    Assume the penalty for a gap is 1 and for a mismatch is 2.
    \begin{enumerate} 
    	\item barbapapa and barbamama 
    	\item barbabravo and barbabright 
    	\item barbabeau and barbabelle 
    	\item barbalala and barbalib
    \end{enumerate}

    \item {[15 points]} 
     Assume we have the following documents in the ``Doc Label: Doc Text'' format (Prof. Hassanpour will cover this topic in the next class):
     \begin{itemize}
     	\item NH: Hampton, Hanover, Keene, Concord, Manchester 
	\item NH: Dartmouth, Manchester, Hanover 
	\item NH: Manchester, Hanover 
	\item VT: Middlebury, Burlington 
	\item VT: Stowe, Rutland, Middlebury, Burlington
     \end{itemize}
     Use this training set and na\"ive Bayes text classification framework with Laplace smoothing to predict the labels the for the following documents:
    \begin{enumerate}
    	\item Dartmouth, Hanover, Burlington
    	\item Middlebury, Dartmouth, Manchester, Burlington
    	\item Stowe, Manchester, Manchester, Burlington
    	\item Keene, Rutland, Stowe
    \end{enumerate}
\end{enumerate}
\end{problem}




\end{document}